---
title: "A01. Assessing abilities: where do we begin?"
author:
- name: Martin Morgan
  affiliation: Roswell Park Comperhensive Cancer Center, Buffalo, NY
date: September 27 2018
output:
  BiocStyle::html_document:
    toc_float: true
vignette: |
  %\VignetteIndexEntry{A01. Assessing abilities: where do we begin?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE"))
)
```

# What are your interests and abilities?

Audiencies are always diverse...

- QUESTION: Divide yourself into percentiles reflecting your expertise
  in biology, statistics, and informatics. For instance, a pure
  biologist might be 90% / 5% / 5%, whereas a statistician with some
  computational skill might be 10% / 60% / 30%.

...and with a wide range of skills

- QUESTION: rate your experience in _R_ by accumulating one point for
  each statement you are comfortable with

    - _R_ is a letter of the alphabet
    - I have used _R_ a little, e.g., to plot data.
    - I have used _R_ scripts
    - I use _R_ regularly for data analysis
    - I have written an _R_ package
    
# How much _R_ do you know?

## Objects and functions are central to _R_

```{r, eval=FALSE}
x <- rnorm(300)
df <- data.frame(x, y = x + rnorm(300))
plot(y ~ x, df)
fit <- lm(y ~ x, df)
anova(fit)
```

QUESTIONS: Understanding _R_

- What fraction of the lines do you understand? Your score will range
  from 0/5 to 5/5.

- How would you add the fitted regression line to the plot?
  
- What does the notation `y ~ x` mean?

- How would you fit a regression line that had to pass through the
  origin?
  
QUESTIONS: _R_ terminology

- There are three objects in the example, `x` ,`df`, and `fit`. Which
  object is an 'atomic' vector? What is the class of each of the other
  objects? What other classes of atomic vector are there?
  
- Which of `rnorm()`, `data.frame()`, `plot()`, `lm()`, and `anova()`
  are plain _R_ functions, and which are 'generic' functions? For the
  generics, what specific method is invoked?

- What class is `y ~ x`? 

QUESTIONS: Navigating _R_'s help

- From the help page for `rnorm()`, `?rnorm`, what other arguments are
  available? What are the default values for those arguments? Is there
  another way within _R_ to find out this information?
  
- How would you find help on the `formula` method for the `plot` generic?

- What methods are available for objects of class `lm`?

## Base _R_ is essential for working with data...

Input

```{r}
dir("~/extdata")
```

```{r}
csv <- read.csv("~/extdata/ALL.csv")
head(csv)
summary(csv)
```

Visualization

- QUESTION: How to visualize the relationship between age and sex?

    ```{r}
    plot(age ~ sex, csv)
    ```
    
Manipulation

- QUESTION: Create a subset containing only females

    ```{r}
    females <- subset(csv, sex == "F")
    ```
    
  What about `csv[csv$sex == "F",]` ?

- QUESTION: create a subset containing 'BCR/ABL' and 'NEG' individuals

    ```{r}
    bcrabl <- subset(csv, mol.biol %in% c("BCR/ABL", "NEG"))
    ```
    
    ```{r}
    summary(bcrabl)
    bcrabl$mol.biol <- factor(bcrabl$mol.biol)
    summary(bcrabl)
    ```
    
## ...but in _R_ there is never only one way: packages

The 'tidyverse' takes a very different approach to working with data

```{r, message=FALSE}
library(tidyverse)
```

Use the tidyverse function `read_csv()` to read our csv file.

```{r}
tbl <- read_csv("~/extdata/ALL.csv")
tbl
```

QUESTIONS: 

- What important differences occur during 'tidy' data input? What
  about in the display of data?

- Use `filter()` to select females wtih BCR/ABL or NEG

    ```{r}
    tbl %>% filter(sex == "F", mol.biol %in% c("BCR/ABL", "NEG"))
    ```
    
- Use `geom_boxplot()` to view filtered data using the 'grammar of
  graphics'

    ```{r}
    tbl %>% filter(!is.na(sex), !is.na(age)) %>%
        ggplot(aes(x = sex, y = age)) + geom_boxplot()
    ```
    
It's useful to reflect on why the tidyverse has become very popular

- A first goal in any analysis is to arrange data into a single
  structure organized along common principles -- 'tidy' data
  represented in a tibble.
- A few 'verbs' enable a lot of work -- `filter()`, `select()`,
  `mutate()`, `group_by()`, `summarize()`.
- Most verbs are 'endomorphic' -- they return the same class of object
  (a tibble) that they take as their first argument
